---
title: "Modeling Social Data Project"
author: "Aarshay Jain | Bindia Kalra | Keerti Agrawal"
date: "May 5, 2017"
output: github_document
---

---
First we read the data and define the names of new variables created, which are:

Explanatory:
* father_involvement
* father_presence
* num_partners
* num_cohab_biof
* num_cohab_anyp

Control:
* mothers_education
* fathers_education
* num_cidi_cases
* hh_income
* kid_punished


```{r}
df <- read.csv('FFChallenge/final_data.csv',header = T, stringsAsFactors = T)
features <- c('father_involvement', 'father_presence', 'num_partners', 'num_cohab_biof', 'num_cohab_anyp','mothers_education', 'fathers_education', 'num_cidi_cases', 'hh_income', 'kid_punished')
outcomes <- c('gpa', 'grit')
head(df[features])
```

# Data Checks:
* missin values
```{r}
apply(df[c(features,outcomes)],2,function(x) sum(is.na(x)))
```
The summary shows everything looks good.
```{r}
summary(df[c(features,outcomes)])
```
Grit and GPA histograms:
```{r,fig.show='hold',fig.align='center'}
hist(df$grit)
hist(df$gpa)
```
Get correlation coefficients:

Though small magnitudes, we see -ve correlation in num_cidi_cases, num_partners and kid_punished and +ve otherwise as expected.
```{r}
apply(df[features[3:length(features)]],2,function(x) cor(x,y=df$gpa))
```
g
