---
title: "Modeling Social Data Project"
author: "Aarshay Jain | Bindia Kalra | Keerti Agrawal"
date: "May 5, 2017"
output: html_document
---

First we read the data and define the names of new variables created, which are:

Explanatory:
* father_involvement
* father_presence
* num_partners
* num_cohab_biof
* num_cohab_anyp

Control:
* mothers_education
* fathers_education
* num_cidi_cases
* hh_income
* kid_punished

```{r}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(stargazer)
```


```{r}
df <- read.csv('FFChallenge/final_data.csv',header = T, stringsAsFactors = F)
features <- c('father_involvement', 'father_presence', 'num_partners', 'num_cohab_biof', 'num_cohab_anyp','mothers_education', 'fathers_education', 'num_cidi_cases', 'hh_income', 'kid_punished')
outcomes <- c('gpa', 'grit')
head(df[features])
```
#Recode factors in a logical sequence:
```{r}
df$father_involvement <- factor(df$father_involvement, levels=c("low","medium","high") )
df$father_presence <- factor(df$father_presence, levels=c("low","medium","high"))
```


# Data Checks:
* missin values
```{r}
apply(df[c(features,outcomes)],2,function(x) sum(is.na(x)))
```
The summary shows everything looks good.
```{r}
summary(df[c(features,outcomes)])
```
Grit and GPA histograms:
```{r,fig.show='hold',fig.align='center'}
# hist(df$grit)
ggplot(data=df)+
  geom_histogram(aes(x=gpa),bins=10, fill="maroon")+
  xlab("GPA at Age 15") +
  ylab("Count")+
  ggtitle("Histogram of GPA") + theme(plot.title = element_text(hjust = 0.5))
```
Get correlation coefficients:

Correaltion between gpa and grit:
```{r}
cor(df$gpa,df$grit)
```

Though small magnitudes, we see -ve correlation in num_cidi_cases, num_partners and kid_punished and +ve otherwise as expected.
```{r}
print('Correlation with GPA:')
apply(df[features[3:length(features)]],2,function(x) cor(x,y=df$gpa))
print('Correlation with Grit:')
apply(df[features[3:length(features)]],2,function(x) cor(x,y=df$grit))
```

Make bar chart for father's involvement and presence:
```{r}
plot1 <- ggplot(data=df) +
  geom_bar(aes(x=father_involvement)) +
  ylab("Count")
 plot2 <- ggplot(data=df) +
  geom_boxplot(aes(x=father_involvement, y=gpa)) +
   ylab("GPA at Age 15")
grid.arrange(plot1, plot2, ncol=2, top="Father's Involvement Feature")
```
Tabulate presence and involvement:
```{r}
table(df$father_involvement, df$father_presence)
```


Make boxplots for father's involvement:
```{r}
plot1 <- ggplot(data=df) +
  geom_bar(aes(x=father_presence)) +
  ylab("Count")
 plot2 <- ggplot(data=df) +
  geom_boxplot(aes(x=father_presence, y=gpa)) +
   ylab("GPA at Age 15")
grid.arrange(plot1, plot2, ncol=2, top="Father's Presence Feature")
```
Though the boxplots are highly overlapping, we can see that father's involvement has a slightly positive impact on GPA. Probably there are many confounding factors which we'll take into consideration in the linear models.

Run a 2-sample t-test:
```{r}
fit <- aov(gpa~father_involvement,data=df)
summary(fit)
fit <- aov(gpa~father_presence,data=df)
summary(fit)
# plot(fit)
# TukeyHSD(fit)
```
Anove gives a highly significant result because it tests that the 3 means are equal. But its a good to compare pairs as well to see which one is actually different. As shown by the boxplot, out hypothesis is confirmed that 

Make boxplots for father's involvement on grit:
```{r}
plot1 <- ggplot(data=df) +
  geom_boxplot(aes(x=father_involvement, y=grit)) 
plot2 <- ggplot(data=df) +
  geom_boxplot(aes(x=father_presence, y=grit)) 
grid.arrange(plot1, plot2, ncol=2)
```

plot others
```{r}
plot1 <- ggplot(data=df) +
  geom_bar(aes(x=num_cohab_biof)) +
  ylab("Count") + xlab("#Cohabiting with Biological Father")
plot2 <- ggplot(data=df) +
  geom_bar(aes(x=num_cohab_anyp)) +
  ylab("Count") + xlab("#Cohabiting with Any Partner")
grid.arrange(plot1, plot2, ncol=2, top="Presence of a Fatherly Figure")
```

```{r}
ggplot(data=df) +
  geom_bar(aes(x=num_partners)) +
  ylab("Count") + xlab("#Number of Relationships With Partners Other than Father")
```


# Linear Models:
we'll start with father's presence and then keep adding new variables in the model:

```{r}
lm1 <- lm(gpa~father_presence,data=df)
# summary(lm1)
lm2 <- lm(gpa~num_cohab_biof+father_presence,data=df)
# summary(lm2)
stargazer(lm1,lm2,type="text",report = "vcp*",intercept.bottom = F, omit.stat=c("adj.rsq","ser","f"))
```
lm2 shows that father_presence is not important in presence of num_cohab_biof which is also a proxy for the same. so we'll just keep one of them.
```{r}
lm3 <- lm(gpa~num_cohab_biof + father_presence  + father_involvement, data=df)
lm4 <- lm(gpa~num_cohab_biof + father_involvement, data=df)
stargazer(lm2,lm3,lm4,type="text",report = "vcp*",intercept.bottom = F, omit.stat=c("adj.rsq","ser","f"))
```
Add more:
```{r}
df$num_cohab_anyp2 <- df$num_cohab_anyp - df$num_cohab_biof
lm5 <- lm(gpa~num_cohab_biof + father_involvement + num_cohab_anyp2, data=df)
stargazer(lm4,lm5,type="text",report = "vcp*",intercept.bottom = F, omit.stat=c("adj.rsq","ser","f"))
```
```{r}
lm6 <- lm(gpa~num_cohab_biof + father_involvement + num_cohab_anyp + num_partners, data=df)
stargazer(lm5,lm6,type="text",report = "vcp*",intercept.bottom = F, omit.stat=c("adj.rsq","ser","f"))
```


```{r}
# lm6 <- lm(gpa~mothers_education+fathers_education+num_cidi_cases+ hh_income+kid_punished,data=df)
# summary(lm6)
lm7 <- lm(gpa~num_cohab_biof + father_involvement + num_cohab_anyp + num_partners +
            mothers_education+fathers_education+num_cidi_cases+ hh_income+kid_punished, data=df)
stargazer(lm6,lm7,type="text",report = "vcp*",intercept.bottom = F, omit.stat=c("adj.rsq","ser","f"))
```
```{r}
lm8 <- lm(gpa~num_cohab_biof + father_involvement + num_cohab_anyp + num_partners +
            mothers_education+fathers_education+ hh_income, data=df)
stargazer(lm7,lm8,type="text",report = "vcp*",intercept.bottom = F, omit.stat=c("adj.rsq","ser","f"))
```

